RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-04 doc/paper.tex
Do Sekcije 4.4.

\begin{itemize}
	\item \fxnote{Marko: Came up to here!}Concerning the exact results in case no noise is utilized in the instance problems, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, begin successful in 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and just 27.14\% of all instances of the benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered benchmark sets. 
	\item   Concerning the exact results when the target noise is presented with a level of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model in more than 42\% cases considering the problems from the both benchmark sets. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the increase in the target noise level, as its performances rapidly drop down -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case the \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} in this order of success, whose percentages of success rate range from $\approx$15--17\%. 
	\item  Concerning the exact results when the presence of noise is significant, that is with a level of 0.01,  our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model for  34.77\% of the all problem instances. The second best is \textsc{Gplearn} which percentage of success is just 12.75\%, followed by \textsc{Dsr} which is successful for just 9.23\% of the all problem instances.

	\item When comparing performances of the algorithms in terms of number of instances for which the algorithm's termination led to an accurate model, that is $R^2 \geq 0.999$, in case when no noise is integrated, our \textsc{Rils}-\textsc{Rols} is successful with the rate of success of 83.38\%. However, in contrast with the exact model percentages comparisons from above, there are a few methods reaching even better percentages of success than ours: \textsc{Mrgp}, and \textsc{Operon}, with 92.69\%, and 86.93\%. On the other hand, as the  exact model percentage rates of these two approaches are rather low (0\% and 16\%, respectively) these indicate the presence of a high level over-fitting  of the later two algorithms. Thus, they are able to obtain a model that is accurate enough on the training data, but most of the produced models are in essence wrong.

   \item  When comparing the performances of the competitors in terms of number of instances for which the  termination led to an accurate enough model, in the presence of the noise level of 0.001, our \textsc{Rils}--\textsc{Rols} method has 78.62\% of a success rate, concerning all considered instances.  The \textsc{Operon} solver only has a better percentage of success among the competitors, which is 83.08\%. However, the over-fitting is conspicuous as this algorithm is rarely able to obtain the model which matches to the exact one (which is successful for 0.31\% cases).

   \item   When comparing the performances of the competitors in terms of number of instances for which the  termination led to an accurate enough model, in the presence of the high--level noise of 0.01, our \textsc{Rils}--\textsc{Rols} method is still highly accurate, delivering  62.92\% of a success rate. Interestingly, it is the best performing in terms of the percentage rate among other competitors. Just for the information, the second best approach is \textsc{Dsr}, with a 16.92\% success rate. It can be concluded that our \textsc{Rils}--\textsc{Rols} method does not pay a big price in presence of a higher noise, like the other methods do, and seem to be very sensible w.r.t. the noise increase. Our \textsc{Rils}--\textsc{Rols} method still delivers high-accuracy models for the noisy input data, but also many of them match to the exact solutions (models).

   \item From the above conclusions, our \textsc{Rils}-\textsc{Rols} algorithm is more robust than the other approaches, as shown by the produced results w.r.t.\ different levels of noise in the input data. Moreover, this is a new state-of-the-art algorithms when one concerns the problem of SR.

@@

\subsection{Scalability of \textsc{Rils}-\textsc{Rols} algorithm}

In this section we study scalability of our method in terms of different noise level in the input data as well as different complexity of the exact solutions (models).

>>>>

\begin{itemize}
	\item Concerning the exact percentages in case no noise is utilized in the input data of the instance problems, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the solution that match to the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, being successful on 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and just 27.14\% of all problem instances from benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered ground-truth benchmark sets. 
	\item   Concerning the exact results in presence of the target noise with a level of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model for more than 42\% cases considering the problem instances from the both benchmark sets. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the increase of the target noise level, as its performances rapidly going down -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case the \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} order decreasingly w.r.t.\ the exact percentages of success, ranging from $\approx$15--17\%. 
	\item  Concerning the exact results when the presence of noise is large, that is with a level of 0.01,  our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model for  34.77\% problem instances from the two benchmark sets. The second best performing is \textsc{Gplearn} which exact percentage of success is just 12.75\%, followed by \textsc{Dsr} which is successful for just 9.23\% problem instances. % over all problem instances.

	\item When comparing performances of the algorithms in terms of number of instances for which the algorithm's termination led to an accurate model, that is $R^2 \geq 0.999$, in case when no noise is integrated, our \textsc{Rils}-\textsc{Rols} is successful with the rate of success of 83.38\%. However, in contrast with the exact model percentages comparisons from above, there are a few methods reaching even better percentages of success than our algorithm: \textsc{Mrgp}, and \textsc{Operon}, with 92.69\%, and 86.93\%, respectively. On the other hand, as the  exact model percentage rates of these two approaches are rather low (0\% and 16\%, respectively) these indicate the presence of a high level over-fitting  of the later two algorithms. Thus, they are able to obtain a model that is accurate enough on the training data, but most of the produced models are in essence wrong.

   \item  When comparing the performances of the competitors in terms of number of instances for which the  termination led to an accurate enough model, in the presence of the noise level of 0.001, our \textsc{Rils}--\textsc{Rols} method has 78.62\% of a success rate, concerning all considered instances.  The \textsc{Operon} solver is the only one that returns a better percentage of success among the competitors, which is 83.08\%. However, the over-fitting is conspicuous as this algorithm is rarely able to obtain the model which matches to the exact one (its exact percentage is equal to 0.31\%).

   \item   When comparing performances of the competitors in terms of number of instances where the termination lead to obtaining an accurate model, in the presence of the high--level noise of 0.01, our \textsc{Rils}--\textsc{Rols} method is still highly accurate, delivering  62.92\% of a success rate. Interestingly, it is the best performing in terms of the percentage rate among other competitors. Just for  the information, the second best approach is \textsc{Dsr}, with 16.92\% of the success rate. It can be concluded that our \textsc{Rils}--\textsc{Rols} method does not pay a big price in presence of a higher noise, like the other methods which are very sensible w.r.t. the noise increase. Our \textsc{Rils}--\textsc{Rols} method still delivers high-accuracy models for the noisy input data, but also many of them fit to the exact solutions (models).

   \item Following the above conclusions, our \textsc{Rils}-\textsc{Rols} algorithm is more robust than the other competitors, which is confirmed by the produced results w.r.t.\ different levels of noise in the input data. Moreover, it can be considered as new state-of-the-art algorithm when one solves the problem of SR.

@@

\subsection{Scalability of \textsc{Rils}-\textsc{Rols} algorithm}
   \fxnote{Marko: Came up to here!}
In this section we study scalability of our method in terms of different noise level in the input data as well as different complexity of the exact solutions (models).
